---
title: "Inferential Analysis - Datasets that assume Normality"
author: "Paolo G. Hilado"
date: "April 30, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Open the data set
```{r}
library(xlsx)
data <- read.xlsx("Experiment2.xlsx", sheetIndex = "Sheet1")
```

## Check the first and last six rows, check the data structure and generate a summary
```{r}
head(data)
tail(data)
str(data)
summary(data)
```

## Clean data by removing the irrelevant first column
```{r}
data <- data[,-1]
```

## Subset data by test administration and group
```{r}
SysPre <- data[data$Group == "system",]$Pre
SysPost <- data[data$Group == "system",]$Post
TradPre <- data[data$Group == "traditional",]$Pre
TradPost <- data[data$Group == "traditional",]$Post
```



## Conduct normality testing for the data for each group and test administration
```{r, message=FALSE}
library(nortest)
nordata <- data.frame(cbind(SysPre, SysPost, TradPre, TradPost))
nortest <- apply(nordata, 2, function(x) ad.test(x))
nortest #Data sets tend to assume Normal Distribution
```

## Generate Probability Density Function Plots
```{r}
par(mfrow=c(2,2))
apply(nordata, 2, function(x) plot(density(x), col = "darkorchid"))
```

## Conduct of test for significant differences in the Pretest of System and Traditional (Unpaired)
```{r, message= FALSE}
library(lawstat)
levene.test(data$Pre, data$Group, location = "mean") #Variances are equal; Student's t-test
t.test(SysPre, TradPre, var.equal = 1) # No significant difference in the pretest ratings between System and Traditional
```

## Conduct of test for significant differences in the Pretest and PostTest for Group System (Paired)
```{r}
t.test(SysPost, SysPre, paired = 1) # There exists a significant difference; it seems that post test ratings tend to be higher than pre test ratings by about 3.88 to 7.37 at 95% confidence interval.
```

## Conduct of test for significant differences in the Pretest and PostTest for Group Traditional (Paired)
```{r}
t.test(TradPost, TradPre, paired = 1) # There exists a significant difference; it seems that post test ratings tend to be higher than pre test ratings by about 4.68 to 6.96 at 95% confidence interval.
```

## Conduct of test for significant differences in the PostTest of System and Traditional (Unpaired)
```{r}
levene.test(data$Post, data$Group, location = "mean") #Variances are equal; Student's t-test
t.test(SysPost, TradPost, var.equal = 1) # No significant difference in the post test ratings between System and Traditional
```

### From the results of the data analysis, it can be observed that the Post test ratings between the use of system and traditional set up are not statistically different thus such may simply be attributed to chance and not necessarily due to the introduction of the treatment. Although a significant increase  can be observed in the ratings for the system group from the pre test to post test, it is to be noted that such is also observed in the traditional group which probably implies that there are other factors that leads to an increase in user ratings apart from the introduction of the treatment. Having no signficant difference in user ratings in the pretests scores of the system and traditional group indicates that the user rating is somewhat similar prior to the introduction of the treatment. 